Predicting Game of the Year (GOTY) 2024: A Bayesian Analysis of Video Game Metrics from Metacritic

### 2024-11-29 other/datasheet/datasheet_template.qmd
add datasheet for dataset

---
title: "Datasheet for '2014-2024 Metacritic video game data'"
author: 
  - Ziqi Zhu
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

@@

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of Australian politicians. We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - TBD
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

@@

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

@@

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

@@

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

@@

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

>>>>

---
title: "Datasheet for '2014-2024 Metacritic video game data on Metacritic'"
author: 
  - Ziqi Zhu
thanks: "Code and data are available at: [https://github.com/zzq20010617/2024TGA-goty-predictions](https://github.com/zzq20010617/2024TGA-goty-predictions)."
date: today
date-format: long
format: pdf
bibliography: references.bib
---

@@

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze factors influencing a video game's likelihood of winning the prestigious Game of the Year (GOTY) award at The Game Awards (TGA). Because I can't found a public dataset that aggregate info for video games from all platforms in recent years. So scrping from website like Metacritic seems to be only choice.

2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset is crated by Ziqi Zhu with scarping of publicly available data sourced from Metacritic [@metacritic2024], a well-known aggregation of reviews for video games, films, and other media.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The dataset is created by Ziqi Zhu, for the purpose of this study only.
4. *Any other comments?*
    - No

@@

	- The instances in the dataset represent video games, with each row corresponding to a single game or downloadable content(DLC) of a game.

@@

	- There are 4207 entries for analyse.

@@

	- The dataset is a sample of video games released between 2014 and 2024 across all platforms. It was specifically curated based on the availability and visibility of games on Metacritic, which primarily lists games with a relatively large influence (e.g., games that have received sufficient attention from users or critics to warrant aggregated reviews). 

@@

	- Each instance in the dataset corresponds to a single video game and includes features(e.g., name, score) that have been preprocessed and engineered from raw data sourced from Metacritic.

@@

	- Yes, goty_status is a binary variable indicating whether a video game won the Game of the Year (GOTY) award at The Game Awards, 1 means game won GOTY for its release year, 0 otherwise.

@@

	- No critical information is missing from individual instances in the dataset after the data cleaning process.

@@

	- No

@@

	- No redundancies after cleaning.

@@

	- The dataset is self-contained.

@@

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data associated with each instance was directly observable, as it was collected by scraping publicly available information from Metacritic.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Python script is used to scrape data from website. The original script is created by Bruno Vieira Ribeiro and posted on Github [@BVRgame] and updated by Ziqi Zhu according to use of this study and the change on Metacritic website, updated version can be find in repo [@myrepo].
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset is scraped from Metacritic, initial sampling is done by Metacritic.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The process is done by creater.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data is collected on November 24, 2024, but it encompasses video games released between 2014 and 2024.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Not applicable
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Data were obtain from third party website Metacritic
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Not applicable
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Not applicable
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Not applicable
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No
12. *Any other comments?*
	- No

@@

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes, the data cleaning process can be find in appendix section in relate paper in Github repo [@myrepo]
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- The raw data can be found in data/01-raw_data directory from Github repo [@myrepo]
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Data was cleaned and labeled by R.
4. *Any other comments?*
	- No

@@

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- No
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No
3. *What (other) tasks could the dataset be used for?*
	- It could be used to analyzing trend of game reviewing system by time and study the correlation between user scores and critic scores.
	4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- Some modification have done based on original collected data, genres with fewer than five games were grouped into an "Other" category to simplify analysis. 
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- Any commercial use is forbidden.
6. *Any other comments?*
	- No

@@

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The dataset will not be distributed to third parties outside of its creator's control. It was developed independently by Ziqi Zhu for academic research purposes and is intended for use only within the scope of this study.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- Not applicable
3. *When will the dataset be distributed?*
	- Not applicable
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Not applicable
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- Not applicable
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- Not applicable
7. *Any other comments?*
	- No

@@

Note that this dataset will not be updated after the completion of this study. This decision was made to avoid repetitive web scraping, which could place unnecessary load on the source platform,

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The dataset will be hosted by the creator, Ziqi Zhu.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Through e-mail of author for this datasheet ziqi.zhu@mail.utoronto.ca
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- No
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- The information collected is at the game level and does not include any personally identifiable information.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- No
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- No external contributions are accepted.
8. *Any other comments?*
	- No

### 2024-11-29 paper/paper.qmd
add datasheet for dataset

I use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse] to process the data. The data were scraped from Metacritic best games from 2014 to 2024 [@metacritic2024], with a modified version of scrape scripts orginally from Bruno Vieira Ribeiro's repo [@BVRgame]. The raw dataset originally have 5968 observations, and after cleaning process @sec-data-clean, there are 4232 left for analyse, 3993 of them are games released between 2014-2023, 239 were released in 2024. I consider use a Bayesian logistic regression model to predict the winner of The Game Award for Game of the Year 2024.

@@

The raw dataset was scraped from Metacritic, a website that aggregates reviews for films, television shows, music albums, and video games. This study focuses on games released between 2014 and 2024. For each game listed on Metacritic, the dataset includes reviews from both critics and users. The data also includes downloadable content (DLC) alongside full games. For example, the action RPG "The Witcher 3: Wild Hunt," released in 2015, and its DLCs "The Witcher 3: Blood and Wine" and "The Witcher 3: Hearts of Stone" are all included. Since The Game Awards (TGA) revised their rules in 2024 to make DLCs eligible for GOTY nominations, these entries are retained in the analysis. In addition to the cleaned version of raw data, I add a column goty_status to mark the games that won GOTY in past years. Preview of first several rows of cleaned dataset ordered by Metascore can be found in @tbl-dataset-preview.

@@

2. Handling Missing Values: Removed rows with missing values in key columns like user_score and user_positivity.
These missing values primarily occurred for games with too few user reviews, as Metacritic does not provide aggregated scores for such entries.

>>>>

I use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse] to process the data. Figures and tables are generated using package **patchwork** [@patchwork], **knitr** [@kabble], packages **broom.mixed** [@broom] and **modelsummary** [@modelsummary] are used to view model summaries, **bayesplot** [@bayesplot] is used to diagnose for model. The data were scraped from Metacritic best games from 2014 to 2024 [@metacritic2024], with a modified Python scrape scripts originally from Bruno Vieira Ribeiro's repo [@BVRgame]. The raw dataset originally have 5968 observations, and after cleaning process @sec-data-clean, there are 4207 left for analyse, 3969 of them are games released between 2014-2023, 239 were released in 2024. I consider use a Bayesian logistic regression model to predict the winner of The Game Award for Game of the Year 2024.

@@

The raw dataset was scraped from Metacritic, a website that aggregates reviews for films, television shows, music albums, and video games. This study focuses on video games released between 2014 and 2024. For each game listed on Metacritic, the dataset includes various features such as release date, genre, the number of reviews from both critics and users, the positive ratio of reviews from critics and users, as well as the aggregated scores from critics and users. The data also includes downloadable content (DLC) alongside full games. For example, the action RPG "The Witcher 3: Wild Hunt," released in 2015, and its DLCs "The Witcher 3: Blood and Wine" and "The Witcher 3: Hearts of Stone" are all included as seperate entries. Since The Game Awards (TGA) revised their rules in 2024 to make DLCs eligible for GOTY nominations, these entries are retained in the analysis. In addition to the cleaned version of raw data, I add a column goty_status to mark the games that won GOTY in past years. Preview of first several rows of cleaned dataset ordered by Metascore can be found in @tbl-dataset-preview.

@@

2. Handling Missing Values, Redundant rows: Removed rows with missing values in key columns like user_score and user_positivity. These missing values primarily occurred for games with too few user reviews, as Metacritic does not provide aggregated scores for such entries. Use "distinct()" to filter out redundant rows, there are 24 redundant rows possibly because of a repeat page in scraping process.
